---
title: "Weight regression"
author: 
  - Noah Siegel
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "documents") })
output:
  pdf_document:
     includes:
        in_header: style.sty
---

```{r environment, include=F}
# Set up global options for nice reports and keeping figures:
knitr::opts_chunk$set(fig.width=14, fig.height=8, fig.align="center", dev = 'png',
                      warning=FALSE, message=FALSE)
```

```{r}
suppressPackageStartupMessages({
  library(plyr)
  library(tidyverse)
  library(sjPlot)
  library(lmerTest)
  library(readxl)
  library(xlsx)
  library(kableExtra)
})
```

```{r directory, include=F}
dir.create('figures/weight/fmt', showWarnings = FALSE, recursive = TRUE)
dir.create('figures/weight/h1n1', showWarnings = FALSE, recursive = TRUE)
dir.create('tables/fmt', showWarnings = FALSE)
dir.create('tables/h1n1', showWarnings = FALSE)
```

```{r echo = FALSE}
data <- read_csv('data/weight_data.csv') %>%
  dplyr::rename(
    Year = year,
    MMU = id,
    Infected = infected,
    Date = date,
    Birthday = birthday
  ) %>%
  mutate(weight_zscore = ave(weight_kg, FUN=scale))
```

```{r echo=F}
# format metadata for publication
data$treatment <- as.character(data$treatment)
data$sex <- as.character(data$sex)
names(data)[names(data)=="treatment"] <- "Treatment"
names(data)[names(data)=="sex"] <- "Sex"
data$Treatment[data$Treatment == "control"] <- "Control"
data$Treatment[data$Treatment == "abx"] <- "Antibiotic"
data$Treatment[data$Treatment == "ABX"] <- "Antibiotic"
data$Treatment[data$Treatment == "abx_fmt"] <- "Antibiotic FMT"
data$Treatment[data$Treatment == "control_H1N1"] <- "Control H1N1"
data$Treatment[data$Treatment == "abx_H1N1"] <- "Antibiotic H1N1"
data$Treatment[data$Treatment == "ABX_H1N1"] <- "Antibiotic H1N1"
```

```{r echo=F}
data$Sex[data$Sex == "M"] <- "Male"
data$Sex[data$Sex == "F"] <- "Female"
data$Treatment <- factor(data$Treatment)
data$Sex <- factor(data$Sex)
```

# Write-up

Using a linear mixed-effects model, it was found that the interaction between treatment
and age (in days) was significantly different when comparing influenza-infected control (CTRL H1N1) and antibiotic (ABX H1N1) animals and control (CTRL) and fecal microbiome transplant (FMT) animals. Sex was not used as a factor in these analyses because it was determined through [analysis of both physiological and gut microbiome data of ABX and CTRL animals](https://github.com/nasiegel88/2021-early-life-gut-microbiome) that males are most heavily impacted by early-life antibiotic treatment.

The original hypothesis was that FMT would reverse the weight changes observed in
ABX and CTRL animals. However, these data suggest that FMT animals weigh more than CTRL
animals. Given that the interaction between age and treatment did not differ between CTRL
and ABX animals, it was surprising to observe significant differences when comparing
FMT and CTRL animals. 

Furthermore, it was hypothesized that influenza infection would exacerbate 
changes observed in ABX and CTRL animals. Our linear mixed-effects model of ABX
and CTRL animals demonstrates a significant difference in the interaction between
sex, age, and treatment, but not age and treatment. With influenza infection and
utilizing age and treatment as factors, we show that influenza-infected animals
that received antibiotics during the first week of life are heavier than animals 
that did not receive antibiotic treatment.

Based on these results, it can be reasoned that FMT does not reverse weight changes
associated with antibiotics since none were observed when sex was not considered.
It is hard to explain what the significant increase in weight of FMT animals could
be since the same effect is observed in ABX H1N1 animals compared to CTRL
H1N1 animals. My first hypothesis is that the increased weight in FMT animals
and ABX H1N1 animals is unrelated. If this hypothesis proves to be valid, there
must be something else going on. 

The best way to address the current discrepancy is to assess whether FMT
reverses the weight changes observed in H1N1 animals. The concern is what significance,
or a lack-there-of, implies. In order to determine what is going on, we need to determine
whether weight changes are good or bad. For example, when an animal is growing and
healthy, positive weight change is good. Conversely, the immune response requires much
energy; thus, negative weight change is often reported in infected animals. What's more, research demonstrates increased weight in response to antibiotics. This begs the question of whether the increased weight observed in ABX H1N1 animals is due to then ABX or the
H1N1.

\newpage

# H1N1 Control vs. Antibiotic

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child('2.1_h1n1.rmd', quiet = TRUE)
cat(res, sep = '\n')
```

\newpage

# FMT vs. Antibiotic

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child('2.2_fmt.rmd', quiet = TRUE)
cat(res, sep = '\n')
```

### Session Information
```{r echo=F}
sessionInfo()
```

